---
title: "Snow Data Assignment: Web Scraping, Functions, and Iteration"
author: "Madeline Bean"
date: "2-7-2022"
output: html_document
---

```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)

```


# Simple web scraping

R can read html using either rvest, xml, or xml2 packages. Here we are going to navigate to the Center for Snow and Avalanche Studies  [Website](https://snowstudies.org/archived-data/) and read a table in. This table contains links to data we want to programatically download for three sites. We don't know much about these sites, but they contain incredibly rich snow, temperature, and precip data. 


## Reading an html 

### Extract CSV links from webpage

```{r}
site_url <- 'https://snowstudies.org/archived-data/'

#Read the web url
webpage <- read_html(site_url)

#See if we can extract tables and get the data that way
tables <- webpage %>%
  html_nodes('table') %>%
  magrittr::extract2(3) %>%
  html_table(fill = TRUE)
#That didn't work, so let's try a different approach

#Extract only weblinks and then the URLs!
links <- webpage %>%
  html_nodes('a') %>%
  .[grepl('24hr',.)] %>%
  html_attr('href')

```

## Data Download

### Download data in a for loop

```{r}

#Grab only the name of the file by splitting out on forward slashes
splits <- str_split_fixed(links,'/',8)

#Keep only the 8th column
dataset <- splits[,8] 

#generate a file list for where the data goes
file_names <- paste0('data',dataset)

for(i in 1:3){
  download.file(links[i],destfile=file_names[i])
}

downloaded <- file.exists(file_names)

evaluate <- !all(downloaded)

```


### Download data in a map

```{r}

#Map version of the same for loop (downloading 3 files)
if(evaluate == T){
  map2(links[1:3],file_names[1:3],download.file)
}else{print('data already downloaded')}

```

## Data read-in 

### Read in just the snow data as a loop

```{r}
#Pattern matching to only keep certain files
snow_files <- file_names %>%
  .[!grepl('SG_24',.)] %>%
  .[!grepl('PTSP',.)]

empty_data <- list()

snow_data <- for(i in 1:length(snow_files)){
  empty_data[[i]] <- read_csv(snow_files[i]) %>%
    select(Year,DOY,Sno_Height_M)
  }

snow_data_full <- do.call('rbind',empty_data)

summary(snow_data_full)
```


### Read in the data as a map function

```{r}

our_snow_reader <- function(file){
  name = str_split_fixed(file,'/',2)[,2] %>%
    gsub('_24hr.csv','',.)
  df <- read_csv(file) %>%
    select(Year,DOY,Sno_Height_M) %>%
    mutate(site = name)
}

snow_data_full <- map_dfr(snow_files,our_snow_reader)

summary(snow_data_full)
```


### Plot snow data

```{r}
snow_yearly <- snow_data_full %>%
  group_by(Year,site) %>%
  summarize(mean_height = mean(Sno_Height_M,na.rm=T))

ggplot(snow_yearly,aes(x=Year,y=mean_height,color=site)) + 
  geom_point() +
  ggthemes::theme_few() + 
  ggthemes::scale_color_few()
```


# Assignment:

1. Extract the meteorological data URLs. Here we want you to use the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets. 

```{r}
links_hw <- webpage %>%
  html_nodes('a') %>%
  .[grepl('forcing',.)] %>%
  html_attr('href')
links_hw
```


2. Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 

```{r}
splits_hw <- str_split_fixed(links_hw, '/', 8)

forcingdataset <- splits_hw[,8] %>% 
  gsub('.txt','',.)

file_names_hw <- paste0('data/', forcingdataset)

for(i in 1:length(file_names_hw)){
  download.file(links_hw[i], destfile=file_names_hw[i])
}

file_names_hw

```

3. Write a custom function to read in the data and append a site column to the data. 

```{r}

# this code grabs the variable names from the metadata pdf file
library(pdftools)
q3_headers <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:26] %>%
  str_trim(side = "left")
q3_headers

q3_reader <- function(file){
  name = str_split_fixed(file, '/', 2)[,2] 
  name2 = str_split_fixed(file, '_', 4)[,2]
  q3test = read.delim(file, header = FALSE, sep ="", col.names = q3_headers, skip = 4) %>%
  select(1:14) %>%
  mutate(site=name2)
}

```

4. Use the `map` function to read in both meteorological files. Display a summary of your tibble.

```{r}
forcing_data_full <- map_dfr(file_names_hw, q3_reader)

summary(forcing_data_full)
```


5. Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.

```{r, warning = FALSE}
q5_yearly <- forcing_data_full %>%
  group_by(year,site) %>%
  summarize(mean_temp_k = mean(air.temp..K.,na.rm=T))
 

ggplot(q5_yearly) + 
  geom_line(aes(x=year,y=mean_temp_k,color=site)) + 
  ggthemes::theme_few() + 
  ggthemes::scale_color_few()
```
What is suspicious about this data is temperature variation at the beginning of the graph, especially since the temperature is in Kelvin. This eludes that temperature was most likely not collected in the earlier years.


6. Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Snow Angel Study Plot?
Hint: https://ggplot2.tidyverse.org/reference/print.ggplot.html

```{r}
forcingmonthyear <- function(forcing_data_full, year){
  monthlytemp<-forcing_data_full %>%
    group_by(month,year,site) %>%
    summarize(monthly_air_temp = mean(air.temp..K.,na.rm=T)) %>%
    filter(yr == year)
  
  plots <- ggplot(monthlytemp, aes(x = month, y = monthly_air_temp, color = site)) +
    geom_line(size=2)+
    facet_wrap(~year)
    labs(title= monthlytemp$year,
         x = "Month",
         y = "Temperature (K)")
  
  print(plots)
}

years <- c(2005,2006,2007,2008,2009,2010)
  
for (yr in years) {
  forcingmonthyear(forcing_data_full, year)
}
  
```
Where both sites follow the same trend for each month, SBSP site is never warmer than SASP site. There are times where are almost the same temperature, but SBSP never exceeds SASP.


Bonus: Make a plot of average daily precipitation by day of year (averaged across all available years). Color each site. 

```{r}
dailytemp<-forcing_data_full %>%
    group_by(day,site) %>%
    summarize(daily_air_temp = mean(air.temp..K.,na.rm=T))

ggplot(dailytemp, aes(x=day, y=daily_air_temp, color=site))+
  geom_line()
```

Bonus #2: Use a function and for loop to create yearly plots of precipitation by day of year. Color each site. 
